\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\title{CSCI 480, Fall 2018\\Math Exercises \#1}
\author{YOUR NAME HERE}
\date{Due date: Friday, October 12, Midnight}
\begin{document}
\maketitle

\begin{itemize}
\item  Copy this file to your own computer and simply add your
  own  name and  answers.
\item   Delete this itemized list from your solutions.
\item Submit both your {\tt hw01.tex} file and your {\tt hw01.pdf}
  file.  Do {\bf NOT} zip or combine them into a single file.
\item
  To make grading easier, please leave the question with
  the answer.
\item
  Show your work and explain your answer.
  If the answer is an integer, I need to know how you
  got it.
\end{itemize}

\begin{enumerate}

\item Explicitly write out the contents of the following set:
  \[ \{X\in \mathcal{P}(\{1,2,3\}) : 2\in X\}\]

  {\bf Answer:}
  

\item Negate the following statement:

  If $x$ is a rational number and $x\neq 0$, then $\tan(x)$
  is not a rational number.

  {\bf Answer:}
  

\item Compute how many 7-digit numbers can be made from the digits
  1,2,3,4,5,6,7 if there is no repetition and the odd digits must
  appear in an unbroken sequence.  Examples: 3571264 or 2415376 or 2467315,
  but not 7234615.


  {\bf Answer:}
  
\item
  This problem concerns 4-card hands dealt off a standard 52-card deck.
  How many 4-card hands are there for which all 4 cards are of different
  suits or all 4 cards are red?


  {\bf Answer:}
  
\end{enumerate}
\end{document}
