\documentclass{article}
\pagestyle{empty}
\usepackage[margin=1in]{geometry}
\usepackage{fancyvrb,amsmath}
\usepackage{multicol}

\newcommand{\set}[1]{\ensuremath{\{#1\}}}
\newcommand{\power}[1]{\ensuremath{\mathcal{P}(#1)}}

\title{CSCI 301, Lab \# 3}
\author{Fall 2018}
\date{}
\begin{document}

\maketitle
%\setlength{\columnsep}{2em}
%\begin{multicols}{2}


\paragraph{Goal:} This lab begins a series of labs that will
build an interpreter for Scheme.  In this lab we get used to
tree recursion as an evaluation strategy, and programming it.
Our programs will be represented by lists, to avoid parsing problems
until later.

\paragraph{Due:} Your program, named {\tt lab03.rkt}, must be submitted to
Canvas before midnight, Monday, May 7.

\paragraph{Unit tests:}
At a minimum, your program must pass the unit tests found in the
file {\tt lab03-test.rkt}.  Place this file in the same folder
as your program, and run it;  there should be no output.  Include
your unit tests in your submission.

\paragraph{Evaluation:}  
The process of evaluating a Scheme expression that consists only of
function calls (procedure applications), for example
{\tt (+  3 x (+ 2 2) (* 2 (+ 4 1)))}, is really very simple.
The process follows three simple rules:
\begin{enumerate}
\item Numbers evaluate to themselves.

  To check whether something
  is a number in scheme, use the {\tt number?} predicate.
\item Symbols, such as {\tt cons}, {\tt +}, and {\tt x},
  are looked up in a table
  called the {\em environment}.  We will use a list for our
  environment.

  To check whether something is a symbol in scheme, use the
  {\tt symbol?} predicate.  (See a pattern here?)
\item Lists are evaluated recursively.  First,
  each element of the list is evaluated, and then the first argument
  (which should be a procedure at this point), is applied to the
  evaluated arguments:

{\tt
  \begin{tabular}{ccccccc}
    ( & + & 3 & x & (+ 2 2) & (* 2 (+ 4 1)) & )\\
    & $\Downarrow$& $\Downarrow$& $\Downarrow$& $\Downarrow$& $\Downarrow$& \\
    ( & \#<procedure:+> & 3 & 5 & 4 & 10 & )\\
    &$\Downarrow$ \\
    &22\\
  \end{tabular}
}
\end{enumerate}

To implement this,  you will write (at least) two procedures:
          {\tt lookup} and {\tt evaluate}.

\paragraph{Lookup:}
The procedure {\tt lookup} is simple.  We represent an environment as
a list of lists.  Each sublist
holds a variable,
and the value of that variable. For example:
\begin{Verbatim}[frame=single]
  (define env (list
                 (list 'x 5)
                 (list '+ +)
                 (list '* *)))
\end{Verbatim}
This would be enough of an environment to evaluate the above
expression.

The {\tt lookup} procedure takes two arguments, a symbol and an
environment, and
\begin{itemize}
\item If the first argument is not a symbol, returns an error.
\item If the symbol is not in the environment, returns an error.
\item Otherwise, returns the value bound to the variable.
\end{itemize}
{\tt lookup} should be written as a simple recursion through the
environment, comparing the provided symbol with the symbol in the
{\tt car} (or {\tt first})
of each variable-value sublist.  If it finds a matching
symbol, the {\tt cadr} (or {\tt second}) thing is returned.
  
\paragraph{Evaluate:}
The {\tt evaluate} procedure takes two arguments, an {\tt expression} to
evaluate, and an {\tt environment}.  It follows the above rules:
\begin{itemize}
\item A number is returned unchanged.
\item For a symbol the return value is looked up in the environment.
\item For a list, each element in the list is evaluated recursively by
  the {\tt evaluate} procedure.  You may
  want to use {\tt map}, but it is not required.
  \begin{itemize}
  \item If the first thing in the evaluated list is not a procedure,
    an error is returned.  (How do you think you check for a procedure
    in Scheme?  Don't overthink it.)
  \item Otherwise, the procedure is applied to the evaluated
    arguments.
  \end{itemize}
\item If the expression is anything else, an error is returned.
\end{itemize}


\end{document}

\end
